---
title       : "Inferential Statistics:"
subtitle    : "Dependent Samples t-Tests"
author      : "William R. Buchanan, Ph.D." 
job         : "Strategic Data Fellow at MDE/Adjunct Prof at JSU"
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides

---
`r opts_chunk$set(message = FALSE)`

## Overview 
* Review
* R stuff
* Model Assumptions
	+ Same subjects
	+ Distribution of the Outcome
* Dependent Samples t-Tests
     + Choosing between dependent and independent samples t-tests
     + Examples
* Next Class

---

## Review
* How many groups of subjects (students) can you compare when using an independent samples t-test?
* What does the p-value from a t-test indicate?
* Explain the 95% Confidence Interval in your own words.
* Are statistically significant differences __always__ practically significant?  Give an example to support your answer
* Give an example of a hypothesis that you could test using a t-test.

---   

```{r, eval = TRUE, echo = TRUE, cache = TRUE}
# Load the foreign data package and the ggplot2 package
library("foreign"); library("ggplot2")

# Store the URL where the data can be accessed
url <- "http://www.paces-consulting.org/stata/JSU/Fall2014/"

# Load either your data or the example data (Substitute your name to get your data)
example <- as.data.frame(read.dta(paste0(url, "JohnDoe-task3.dta")))
```

---   

## R Stuff
* As you'll see soon enough, dependent samples t-tests do things a bit differently than
independent samples t-tests, and in order to fully appreciate the assumptions, we need to 
be able to do a bit of data manipulation
* You've done basic arithmetic in R during the DataCamp assignment, and now we're going to 
compare that with a little bit of data management to help you look at/assess the assumptions 
of the model you'll end up fitting to your data
* Remember that you can refer to a variable in a dataframe using these types of references
     + `example$mthscore2`
     + `example["rlascore1"]`
     + `example[, 14]`
* In the first example `example` is the name of your data object, `$` is a symbol used to reference a column within that data object, and `mthscore2` is the name of the variable you're interested in
* In the second example `example` is still the name of your data object and `["rlascore1"]` is the name of the variable you're interested in
* In the last example, `example` is still the name of your data object and `[, 14]` is the position - or column number - for the variable you are interested in

---   

* You should remember that you can use objects for basic arithmetic:
     + `x <- 5; y <- 3; x - y` (you should see `[1] 2` if you do that in R)
* You should also remember from the graph examples that you can _recursively_ create objects
     + In this context, remember the cases in the examples/class where we assigned a result to an 
     object that included itself like `x <- 5; x <- x - 3`
* We can combine those ideas to create new variables that contain results like this as well

---   

```{r, eval = TRUE, echo = TRUE, tidy = TRUE}
# Create a new variable containing the difference between math scores
example$mthdiff <- example$mthscore2 - example$mthscore1

# Create a new variable containing the difference between reading scores
example$rladiff <- example$rlascore2 - example$rlascore1

# Look at summaries of these variables
summary(example[c("rladiff", "mthdiff")])
```

---   

# Distribution of the outcome
* The data must still be measured on the intervallic/ratio scale
* The difference between __independent__ and __dependent__ t-tests assumptions is :
     + Independent samples t-tests assume the distribution of the _outcome variable_ is approximately 
     Gaussian (normal; looks like a bell curve)
     + Dependent samples t-tests assume that the __difference__ between the two measures being 
     compared are approximately Gaussian (normal; looks like a bell curve)
* The way that you can assess this, is by doing what we did above (e.g., creating a new variable with 
the difference between the two meausres) and graphing that distribution

---   

```{r, echo = TRUE, eval = TRUE, fig.align = "center", warning = FALSE, fig.height = 6, fig.width = 12}
# Create a graph of the distribution of the difference in the scores
ggplot(data = example, aes(x = mthdiff)) + geom_density() + #
ggtitle("Distribution of the Difference in Math Scores: \nPrior vs Current Year") + #
xlab("Difference Score") + ylab("Density of Observations")
```

---

```{r, echo = TRUE, eval = TRUE, fig.align = "center", warning = FALSE, fig.height = 6, fig.width = 12}
# Create a graph of the distribution of the difference in the scores
ggplot(data = example, aes(x = rladiff)) + geom_density() + #
ggtitle("Distribution of the Difference in Reading Scores: \nPrior vs Current Year") + #
xlab("Difference Score") + ylab("Density of Observations")
```

---

# Same Subjects
* For __independent__ samples t-tests the students can only appear in one group or the other
* For __dependent__ samples t-tests you are comparing the students to themselves across two measures
* So for a __dependent__ or __paired__ samples t-test, there is no grouping variable because you are 
comparing the difference in the means of the same students (e.g., comparing how students in your 
class did on test 1 with how they did on test 2)

---

## Dependent Samples t-Tests
* Last week, when we discussed independent samples t-tests, the discussion focused on two distinct 
groups of subjects (students in our example data) that did not overlap (e.g., a student could only 
belong to a single group)
     + _independent samples_ is actually telling us that the two groups should be mutually exclusive
     (e.g., cannot belong to both groups) and should be independent of one another (e.g., you 
     wouldn't compare a subset of a group with the group as a whole [e.g., you wouldn't compare the
     mean of Hispanic/Latino males to the mean of all Hispanics/Latinos])
* __Dependent samples__ or __paired samples__ t-tests have a very different assumption about the 
subjects, which is that you are comparing the same subjects across two measures (e.g., comparing 
the students' scores from last year with the current year)
     + _dependent_ or _paired_ samples implies that the measures are paired by subjected (e.g., the
     outcome variables are each paired with the same student) and/or that the measures are dependent 
     on time (e.g., the same measurement is used at time 1 and time2)
     + This may sound like there isn't a difference, but _paired_ implied that you can test the 
     difference in the means of reading and math tests for the same students (e.g., see if there is 
     a difference in the means of the two scores)
* Give an example of when you may want to use a dependent/paired samples t-test

---   

## Choosing between dependent and independent samples t-tests?
* If you are interested in comparing two distinct groups of observations use the __independent__ 
samples t-test
     + You want to compare the scores of students getting free lunch vs students not getting free 
     lucn on the same test
     + You want to compare the heights of boys vs girls
* If you are interested in comparing the same students over time, or the same students on two 
different measures, use a __dependent__ samples t-test
     + You want to compare whether there is a difference in reading and math scores for the students
     in your classroom
     + You want to compare whether or not the difference between student scores on the first 9-week 
     exam is different from their scores on the second 9-week exam (e.g., did the scores go up down 
     or stay the same on average)
* Remember that although it is possible to do some of these things, it may not make practical or 
theorical sense to do so
* When deciding if you should/shouldn't do something for a t-test ask yourself if it is something 
that you think would be helpful to you and/or your colleagues to know about the student outcomes

---    

## Examples
```{r, echo = TRUE, eval = TRUE, tidy = TRUE}
# Test first hypothesis example
t.test(example$mthscore2, example$mthscore1, paired = TRUE)
```

---   

## Next Class
* Simple correlations 
* Use the task3 dataset for the task4 assignment
* There may be a new data set for the final so be sure that you are comfortable finding out how to 
label your graphs using information from the dataset (in previous weeks' slides) and are comfortable 
substituting different variables into the examples















